Measuring neutron star tidal deformability with Advanced LIGO: a Bayesian analysis of mixed binary observations

### 2016-05-26 Paper/paper.tex
text in s2a

In this section we summarize various technical aspects of this paper. The model
used to obtain tidal waveforms is described in Sec.~\ref{s2:waveforms}. The 
process of inferring source parameters from a GW signal is summarized in 
Sec.~\ref{s2:bayesian}. The process of generating and combining multiple events
is instead detailed in the corresponding results section,
Sec.~\ref{s2:astro_multiple}.

@@

In Ref.~\cite{Lackey:2013axa}, the authors presented a suite of $134$ numerical
relativity simulations of neutron stars inspiraling into spinning black holes. Amongst
the varied quantities were (a) the equation of state for the NS, (b) NS mass, 
(c) mass-ratio and (d) black hole spins, taken as aligned with the orbit.
A total of $21$ two-parameter EoSs were used for these simulations, which also 
had $q\in[2, 5]$, $\chibh\in[-0.5, +0.75]$, and 
$m_\mathrm{NS}\in[1.20, 1.45]M_\odot$. Based on this suite of simulations, the
authors calibrated a frequency-domain waveform model for the inspiral and merger phasing of
NSBH coalescences. The GW strain predicted by this model (henceforth ``LEA''),
can be written as
% 
\begin{equation}
 \tilde{h}_\mathrm{NSBH}(f, \vec{\theta}, \lambdans) = \tilde{h}_\mathrm{BBH}(f, \vec{\theta})\,A(f, \vec{\theta}, \lambdans)\,e^{\ii \Delta\Phi(f, \vec{\theta}, \lambdans)},
\end{equation}
where $\vec{\theta}:=\{\mbh, \mns, \chibh,\chins=0\}$ are the intrinsic binary
parameters. The first term on the right hand side is an underlying
BBH waveform model that is used as a base. In the original construction of LEA,
this was taken to be the SEOBNRv1 model~\cite{Taracchini:2012} of the 
Effective-one-body (EOB) family~\cite{Buonanno99}. The second term changes the 
amplitude of the BBH model
to match that of an NSBH merger of otherwise identical parameters, with NS-matter
effects parametrized by $\lambdans$. During inspiral this term is set to
unity, but is a sensitive function of $\lambdans$ close to merger. The third 
term does the same as the second, but to the waveform phasing. During inspiral,
$\Delta\Phi(f, \vec{\theta}, \lambdans)$ is set to the PN tidal phasing corrections,
at the leading and next-to-leading orders~\cite{Vines2011}. Close to merger,
additional phenomenological terms are added and calibrated to the $134$ available
NR simulations.

In this paper we use LEA for our signal and template modeling, but switch the 
underlying BBH model to SEOBNRv2~\cite{Taracchini:2013rva},
using its reduced-order frequency-domain version~\cite{Purrer:2015tud}. We
expect this change to make our conclusions more robust because: (a) the 
SEOBNRv2 model is more accurate~\cite{Kumar:2015tha,Kumar:2016dhh}, and (b)
the differences between the two EOB models are caused by the
inaccuracies of SEOBNRv1 during the {\it inspiral} phase, many orbits before 
merger~\cite{Kumar:2015tha}.
Since LEA only augments the inspiral phasing with PN tidal terms, our
change in the underlying BBH model does not change LEA's construction, but
increases the overall model accuracy during inspiral. Therefore the model used 
here (``LEA+'') is a more robust version of LEA.

>>>>

% In this section we summarize various technical aspects of this paper. The model
% used to obtain tidal waveforms is described in Sec.~\ref{s2:waveforms}. The 
% process of inferring source parameters from a GW signal is summarized in 
% Sec.~\ref{s2:bayesian}. The process of generating and combining multiple events
% is instead detailed in the corresponding results section,
% Sec.~\ref{s2:astro_multiple}.

@@

% \red{
% Begin with introducing sentence:
% }

Lackey et al.~(LEA)~\cite{Lackey:2013axa} developed a complete inspiral-merger
waveform model for disrupting NSBHs. Theirs is a frequency-domain
phenomenological model that includes the effect of BH and NS masses and spins
$\{\mbh, \chibh, \mns\}\equiv\vec{\theta}$ and NS tidal deformability
$\lambdans$. It was calibrated to a suite of $134$ numerical relativity (NR)
simulations of NSs inspiraling into spinning BHs. The parameter space these
simulations span, which we assume to be the domain of validity for the LEA
model, includes NS masses $1.2M_\odot\leq\mns\leq 1.45M_\odot$,
mass-ratios $2\leq q\leq 5$, and BH spins $-0.5\leq\chibh\leq+0.75$.
They also sample a total of $21$ two-parameter nuclear EoSs to cover the
spectrum of NS deformability.
% 
% Amongst the varied quantities were (a) the
% equation of state for the NS, (b) NS mass, (c) mass-ratio and (d) black hole
% spins, taken as aligned with the orbit.
% A total of $21$ two-parameter EoSs were used for these simulations, which also 
% had $q\in[2, 5]$, $\chibh\in[-0.5, +0.75]$, and 
% $m_\mathrm{NS}\in[1.20, 1.45]M_\odot$. Based on this suite of simulations, the
% authors calibrated a frequency-domain waveform model for the inspiral and merger phasing of
% NSBH coalescences.
% 
The GW strain $\tilde{h}(f)$ per the LEA model can be written as
\begin{equation}
 \tilde{h}_\mathrm{NSBH}(f, \vec{\theta}, \lambdans) = \tilde{h}_\mathrm{BBH}(f, \vec{\theta})\,A(f, \vec{\theta}, \lambdans)\,e^{\ii \Delta\Phi(f, \vec{\theta}, \lambdans)},
\end{equation}
with NS spin $\chins=0$ identically. Here, $\tilde{h}_\mathrm{BBH}$ is
an underlying BBH waveform model. In the original LEA model,
this was taken to be the SEOBNRv1
model~\cite{Taracchini:2012} of the Effective-one-body (EOB)
family~\cite{Buonanno99}. The factor $A(\cdot)$ adjusts
the amplitude of the BBH model to match that of an NSBH merger of otherwise
identical parameters, with NS-matter effects parametrized by $\lambdans$.
During early inspiral this term is set to
unity, but is a sensitive function of $\lambdans$ close to merger. The term with
$\Delta\Phi$ corrects the waveform phasing. During inspiral,
$\Delta\Phi$ is set to the PN tidal phasing corrections,
at the leading and next-to-leading orders~\cite{Vines2011}. Close to merger,
additional phenomenological terms are needed. Both $A$ and $\Delta\Phi$ are
calibrated to all $134$ available NR simulations.

In this paper we use LEA for our signal and template modeling, but switch the 
underlying BBH model to SEOBNRv2 (and refer to it as enhanced-LEA or
``LEA+'')~\cite{Taracchini:2013rva}. We using the reduced-order
frequency-domain version of SEOBNRv2, which has the additional benefit of
reducing computational cost~\cite{Purrer:2015tud}. We expect this change
(LEA$\rightarrow$LEA+) to make our conclusions more robust because: (a) the 
SEOBNRv2 model is more accurate~\cite{Kumar:2015tha,Kumar:2016dhh}, and (b)
the differences between the two EOB models are caused by the
inaccuracies of SEOBNRv1 during the {\it inspiral} phase, many orbits before 
merger~\cite{Kumar:2015tha}.
Since LEA only augments inspiral phasing with PN tidal terms, our
change in the underlying BBH model does not change LEA's construction, {\it and}
increases the overall model accuracy during inspiral.
% We will refer to this enhanced-LEA model henceforth as ``LEA+''.
